Regression

### 2017-12-08 paper/R/04_regression.Rmd
new

```{r Setup, include=FALSE}
suppressPackageStartupMessages({
  library(dplyr)       # Data manipulation
  library(stringr)     # String manipulation
  library(lubridate)   # Date and time manipulation
  library(purrr)       # Functional programming
  library(tidyr)       # Reshaping
  library(magrittr)    # Advanced piping
  library(pushoverr)   # Pushover notifications
  library(doMC)        # Parallel Computing
  library(readr)       # Importing data
  library(tibble)      # Better data frames
  library(data.table)
  library(pscl)
  library(boot)

  library(ggplot2)     # Static data visualization
  library(scales)      # Scales
  library(viridis)     # Viridis color scales

@@

theme_set(
  theme_bw(base_family = "Roboto", base_size = 14) +
    theme(
      plot.title = element_text(face = "bold", size = 14,
                                margin = margin(0, 0, 4, 0, "pt")),
      plot.subtitle = element_text(size = 12),
      plot.caption = element_text(size = 6, hjust = 0),
      axis.title = element_text(size = 10),
      panel.border = element_blank()
    )
)

rm(list=ls())
```

## Load Data
```{r caching, echo = FALSE}
load("../output/btw_combined.Rda")
btw2 <- btw

load("../output/models/STM 40 .Rda")
#load("../output/btw_combined.Rda")
knitr::opts_chunk$set(eval = TRUE, warning = FALSE, message = FALSE)

# Set color
library(RColorBrewer) 
col <- brewer.pal(6,"Dark2")
```

## Prepare Data
```{r Extract wtp and dtp}
# Word-topic probabilities
stmOut %>% tidy("beta") %>% filter(!is.na(topic)) -> posts.wtp
# Document-topic probabilities
stmOut %>% tidy("gamma") -> posts.dtp
```

```{r}
top_topics <-
  posts.dtp %>% 
  group_by(document) %>%
  mutate(therank = rank(-gamma, ties.method = "random")) %>%
  filter(therank %in% 1) %>%
  select(- therank)
```

We classify each document based on which topic has the highest probability. To estimate the true effect of a topic, we delete documents where the highest topic probability is smaller than 0.5, as topic assignment is not ambiguous. This left us with 7740 documents.

```{r}
btw.reg <- 
  btw %>%
  left_join(., btw2 %>%
              select(title.text, shares, url),
            by = "title.text")
```

```{r}
# filter documents with low posterior
btw.reg <- 
  btw.reg %>%
  mutate(document = as.numeric(articleID)) %>%
  left_join(., top_topics, by="document") %>%
  filter(gamma > 0.5)

# generate new variables and reduce df
btw.reg %>%
  mutate(topic = factor(topic),
         site = factor(site),
         shares.dummy = factor(ifelse(shares==0,0,1))) %>%
  select(site, shares, shares.dummy, date, text_length, month, 
         document, topic, gamma) -> btw.reg
```

## Description of Data
```{r}
summary(btw.reg)
```

The proportion of zeros in the independent variable (number of Facebook shares) is about 65.6\%. One way to model this type of situation is to assume that the data come from a mixture of two populations, one where the counts is always zero, and another where the count has a Poisson distribution with mean $\mu$. In this model zero counts can come from either population, while positive counts come only from the second one.

The distribution of the outcome can then be modeled in terms of two parameters, $\pi$ the probability of 'always zero', and $\mu$, the mean number of publications for those not in the 'always zero' group. A natural way to introduce covariates is to model the logit of the probability $\pi$ of always zero and the log of the mean $\mu$ for those not in the always zero class.

## Analysis
### Zero-inflated Poisson regression
```{r}
# filter NAs
btw.reg %>%
  filter(!is.na(shares)) -> btw.reg
```

```{r Zero-inflated Poisson}
library(pscl)
m1 <- zeroinfl(shares ~ topic + site | topic + site + text_length, data = btw.reg)
summary(m1)
```
Looking at the inflate equation we see that some topics are significant predictors of being in the 'always zero' class.

@@

```{r}
p1 <- glm(shares ~ topic, family = poisson, data = btw.reg)
summary(p1)
```

@@

The predicted values are plotted in black. They are connected to the actual data points through a line and high residuals (in abolsute terms) are made more red on actual values. This helps to identify non-linearity in the data. We can see that there is more red for extreme values for spiegel.de as well as for the topics 12, 21 and 9. 
```{r message=FALSE, warning=FALSE}
btw.reg %>%
  select(site, shares, topic, residuals, predicted) %>%
  gather(key="iv", value="x", -shares, -predicted, -residuals) %>%
  ggplot(aes(x=x, y=shares)) +
  geom_segment(aes(xend=x, yend=predicted), alpha=.2) +
  geom_point(aes(color=residuals)) +
  scale_color_gradient2(low="black", mid="blue", high="red") +
  guides(color=FALSE) +
  geom_point(aes(y=predicted), shape=1) +
  facet_grid(~iv, scales="free") +
  theme_bw()
```

>>>>

```{r}
suppressPackageStartupMessages({
  library(dplyr)       # Data manipulation
  library(stringr)     # String manipulation
  library(lubridate)   # Date and time manipulation
  library(purrr)       # Functional programming
  library(tidyr)       # Reshaping
  library(magrittr)    # Advanced piping
  library(pushoverr)   # Pushover notifications
  library(doMC)        # Parallel Computing
  library(readr)       # Importing data
  library(tibble)      # Better data frames
  library(data.table)
  library(pscl)
  library(boot)
  library(stm)
  library(readxl)

  library(ggplot2)     # Static data visualization
  library(ggpmisc)
  library(RColorBrewer) 
  library(ggrepel)
  library(plotly)
  library(scales)      # Scales
  library(viridis)     # Viridis color scales

@@

theme_set(
  theme_bw(base_family = "Roboto", base_size = 8) +
    theme(
      plot.title = element_text(face = "bold", size = 14,
                                margin = margin(0, 0, 4, 0, "pt")),
      plot.subtitle = element_text(size = 12),
      plot.caption = element_text(size = 6, hjust = 0),
      axis.title = element_text(size = 10),
      panel.border = element_blank()
    )
)

#rm(list=ls())
col <- brewer.pal(7,"Dark2")
```

# Load Data
```{r caching, echo = FALSE}
#load("../output/btw_combined.Rda")
load("../output/models/STM 28 .Rda")
```

load data "Unique User"
```{r}
## Daily (06-2017 - 12-2017)
unique_user1 <- read_excel("../data/uniqueUser_1.xls")
unique_user2 <- read_excel("../data/uniqueUser_2.xls")

uu_df <- rbind(unique_user1,unique_user2) %>%
  rename(uniqueUser = `Unique User Tag Mio.`,
         site = Dachmarke,
         date = Tag) %>%
  select(date, site, uniqueUser)

rm(unique_user1, unique_user2)
```

# Description of Data

### Unique User 
```{r}
p <- uu_df %>%
  filter(date > as.Date("2017-05-31")) %>%
  ggplot(aes(date, uniqueUser, color=site)) +
  geom_line() +
  geom_vline(aes(xintercept=as.POSIXct("2017-09-24")),
             linetype = 2, color="grey50") +
  scale_color_manual(values = col[c(1,2,3,4,6,7)]) +
  theme(
      #legend.position   = "none",
      legend.text       =  element_text(size=6),
      plot.background   = element_rect("#fafafa", "#fafafa"),
      panel.background  = element_rect("#fafafa"),
      axis.title.x      = element_blank(),
      axis.text.y       = element_text(size = 8)
    ) +
  labs(x="", y="", color="", title = "", 
      caption = "
  Source: AGOF (www.agof.de)
      " )

#ggplotly(p)
ggsave(plot= p,
filename = "../figs/uniqueUser.png", device = "png",
width = 6, height = 4,
        dpi = 600
)
```

### Distribution of articles
```{r message=FALSE, warning=FALSE}
ggsave({
  btw %>%
  ggplot(aes(site)) +
  geom_bar(fill=col, alpha= .8) +
  labs(x="", y="Count") +
  theme(
      legend.position   = "none",
      plot.background   = element_rect("#fafafa", "#fafafa"),
      plot.title        = element_text(size = 18),
      plot.subtitle     = element_text(size = 8),
      plot.caption      = element_text(size = 7),
      panel.background  = element_rect("#fafafa"),
      axis.title.x      = element_blank(),
      axis.text.y       = element_text(size = 8)
    )

},
filename = "../figs/bar.png", device = "png", 
width = 6, height = 4,
        dpi = 600)
```

```{r message=FALSE, warning=FALSE}
ggsave({
  btw %>%
  group_by(date, site) %>%
  summarise(obs = n()) %>%
  ggplot(aes(date, obs, color = site)) +
  geom_line() +
  #geom_vline(aes(xintercept=as.Date("2017-09-24")),
  #           linetype = 2, color="grey40") +
  #geom_vline(aes(xintercept=as.Date("2017-11-20")),
  #               linetype = 2, color="grey40") +
  scale_color_manual(values = col) +
  labs(x="", y="Count",color="") +
  scale_x_date(breaks = date_breaks("1 month"), labels=date_format("%B", tz="CET")) +
  theme(
      #legend.position   = "none",
      plot.background   = element_rect("#fafafa", "#fafafa"),
      plot.title        = element_text(size = 18),
      plot.subtitle     = element_text(size = 8),
      plot.caption      = element_text(size = 7),
      #panel.grid        = element_blank(),
      panel.background  = element_rect("#fafafa"),
      axis.title.x      = element_blank(),
      axis.text.y       = element_text(size = 8)
    )
},
filename = "../figs/timeline.png", device = "png",width = 7, height = 4,
dpi = 600
)

```

# Model Results
```{r Extract wtp and dtp}
# Word-topic probabilities
stmOut %>% tidy("beta") %>% filter(!is.na(topic)) -> posts.wtp
# Document-topic probabilities
stmOut %>% tidy("gamma") -> posts.dtp
```

## Topic Labeling
```{r Label Topics}
label <- labelTopics(stmOut, n=3)

prob <- as.data.frame(label$prob, stringsAsFactors = F)
frex <- as.data.frame(label$frex, stringsAsFactors = F)
lift <- as.data.frame(label$lift, stringsAsFactors = F)
score <- as.data.frame(label$score, stringsAsFactors = F)

topicLabel <- prob %>% 
  transmute(topic = rownames(.),
            topic_name = paste(prob$V1,prob$V2,prob$V3,score$V1,score$V2,score$V3, sep=","),
            prob = paste(prob$V1,prob$V2,prob$V3, sep=","),
            frex = paste(frex$V1,frex$V2,frex$V3, sep=","),
            lift = paste(lift$V1,lift$V2,lift$V3, sep=","),
            score = paste(score$V1,score$V2,score$V3, sep=","))

topicLabel$topic_name <- vapply(lapply(strsplit(topicLabel$topic_name, ","), unique), paste, character(1L), collapse = " ")
rm(prob, frex, lift, score)
```

```{r}
top_topics <-
  posts.dtp %>% 
  group_by(document) %>%
  mutate(therank = rank(-gamma, ties.method = "random")) %>%
  filter(therank %in% 1) %>%
  select(- therank)

# Add Topic to origian DF
btw %>%
  mutate(document = articleID) %>%
  inner_join(.,top_topics, by="document") %>%
  ## Combine with Topic label
  left_join(., topicLabel %>%
              select(topic_name, topic) %>%
              mutate(topic = as.numeric(topic)), by="topic") -> btw
```

## Topic Proportion
```{r}
png("../figs/topic_proportion.png")
plot(stmOut, type = "summary", xlim = c(0,.16), n = 5, 
     custom.labels = topicLabel$topic_name, 
     family = "Roboto", main = "", text.cex = 0.7)
dev.off()
```

## Topic Correlation
We can correlate the topic by the metadata available. The column means of theta, grouped by category is calculated.
```{r}
theta <- as.data.frame(stmOut$theta)
names(theta) <- as.character(1:28)

theta2 <- theta %>% mutate(document = as.numeric(rownames(theta))) %>%
  left_join(., btw %>% select(site, document),
            by="document")

theta.mean.by <- by(theta2[,1:28], theta2$site, colMeans)
theta.mean <- do.call("rbind", theta.mean.by)

library(corrplot)
c <- cor(theta.mean)
corrplot(c, method = "circle")
```

## Sample Articles
We classify each document based on which topic has the highest probability.

```{r Document classification}
set.seed(9272)
posts_classification.sdt <-
  posts.dtp %>% 
  group_by(articleID = document) %>% 
  summarise(topic = min(topic[gamma == max(gamma)]), gamma = max(gamma)) %>% 
  ungroup() %>%
  inner_join(btw %>% select(title, title_text, articleID, topic_name),
              by = "articleID") %>%
  #inner_join(topics_mapping.dt, by = "topic") %>%
  mutate(topic_title = 
           paste0("Topic ", formatC(topic, flag = "0", width = 2), 
                  " - ", topic_name)) %>%
  group_by(topic_name) %>% 
  top_n(300, gamma) %>%
  sample_n(7) %>% 
  mutate(row = row_number()) %>% 
  ungroup()

```

```{r Document classification chart, echo=TRUE}

ggsave(
  plot = {
    posts_classification.sdt %>% 
      ggplot(aes(x = 0, y = row, color = factor(topic, levels = 1:28))) +
      facet_wrap(~topic_title, ncol = 4) +
      geom_text(
        aes(label = title),
        size = 3, hjust = 0, family = "Roboto"
      ) +
      geom_point(aes(x = -0.025), size = 0.5) +
      scale_x_continuous(limits = c(-0.05, 1), expand = c(0, 0)) +
      scale_color_manual(
        values = rainbow(28) %>% 
          adjustcolor(red.f = 0.6, green.f = 0.6, blue.f = 0.6)
      ) +
      theme(
        axis.title       = element_blank(),
        axis.text        = element_blank(),
        axis.ticks       = element_blank(),
        panel.grid       = element_blank(),
        strip.background = element_blank(),
        panel.border     = element_blank(),
        strip.text       = element_text(face = "bold", hjust = 0, size = 7),
        plot.background  = element_rect(fill = "#FAFAFA"),
        legend.position  = "none"
      ) 
  },
  device = "png",
  filename = "../figs/topic-classification-sample.png",
  dpi = 600, height = 10, width = 16
)
```

## Estimating the Effect of Covariates

### Differences in Word-Topic Distribution
How do different news sources describe the same topic with different vocabulary

```{r Top Words by topic and news source}
btw %>% 
  unnest_tokens(word, text_cleaned, token = "words") %>%
  filter(str_length(word) >= 3) %>%
  count(site, topic, word) %>%
  ungroup() ->
  posts_tokenized.dt

posts_tokenized.dt %>%
  group_by(site, topic) %>%
  dplyr::arrange(desc(n)) %>%
  top_n(18) %>%
  # concatenate words
  group_by(site, topic) %>%
  summarise(top_words = paste(word, collapse=", ")) %>%
  mutate(top_words = gsub("([^,]+,[^,]+,[^,]+),","\\1\n", top_words)) %>%
  left_join(.,topicLabel %>%
              mutate(topic=as.numeric(topic),
                     topic_name = gsub("([^\\s]+,[^\\s]+),","\\1\n", topic_name)) %>%
              select(topic, topic_name),
            by="topic") ->
  top_words.dtf
```

```{r}
ggsave(
  plot = {
    top_words.dtf %>% 
  filter(topic %in% c(5,6,7,12,15,16)) %>%
  mutate(row=row_number()) %>%
      ggplot(aes(x = topic_name, y=site, color=site)) +
      facet_wrap(~topic_name, ncol=6) +
      geom_text(aes(x =topic_name, y=site, label=top_words),
                hjust="inward",
                size = 2, family = "Roboto"
      ) +
      theme(
        axis.title       = element_blank(),
        axis.text        = element_blank(),
        axis.ticks       = element_blank(),
        panel.grid       = element_blank(),
        strip.background = element_blank(),
        panel.border     = element_blank(),
        strip.text       = element_text(hjust = 0.5, size = 6),
        plot.background  = element_rect(fill = "#FAFAFA"),
        legend.title     = element_text(size=7),
        legend.text      = element_text(size=7)
      ) 
      #guides(colour = guide_legend(override.aes = list(size=1)))
  },
  device = "png",
  filename = "../figs/top_words_news.png",
  dpi = 600, height = 6, width = 12
)
```

### Difference in topic prevalence
The estimateEffect() function explores how prevalence of topics varies across documents according to document covariates (metadata). First, users must specify the variable that they wish to use for calculating an effect. If there are multiple variables specified in estimateEffect(), then all other variables are held at their sample median. These parameters include the expected proportion of a document that belongs to a topic as a function of a covariate, or a first difference type estimate, where topic prevalence for a particular topic is contrasted for two groups.

```{r}
prep <- estimateEffect(1:28 ~site + s(month),
                       stmOut, meta=out$meta, uncertainty = "Global")
```

```{r}
summary(prep)
```

Expected difference in topic probability by news source (with 95% Confidence Intervals). 
parameter "covariate": String of the name of the main covariate of interest. Must be enclosed in quotes. All other covariates within the formula specified in estimateEffect will be kept at their median.
```{r}
for (i in 1:28){
  png(paste0("../figs/estimate_effect",i,".png"))
  plot(prep, "site", method = "pointestimate", topics = i,
       labeltype = "custom", custom.labels = unique(prep$data$site),
       ylab = "", xlab = "Mean topic proportion in corpus",
       main = paste0("Topic ",i,": ",topicLabel$topic_name[i]),
       xlim = c(-.05,0.15))
  dev.off()
}
```

```{r}
for (i in 1:25){
  plot(prep, "month", method = "continuous", topics = i,
       printlegend = F,
       ylab = "", xlab = "Mean topic proportion in corpus",
       main = paste0("Topic ",i,": ",topicLabel$topic_name[i]))
}

```

```{r}
p<- posts.dtp %>%
  left_join(., btw %>% select(document, topic_name,
                              site, month),
            by="document")  %>%
  mutate(site = factor(site),
         topic = factor(topic)) %>%
  filter(month!=5) %>%
  ggplot(aes(weight=gamma, x=topic, fill=topic, text=topic_name)) +
  geom_bar() +
  coord_flip() +
  facet_grid(site~month) +
  guides(fill=FALSE) +
  xlab("Proportion")

ggplotly(p)
```

# Regression

What is the distribution of the gamma value?
```{r}
ggsave(plot={
  btw %>%
  ggplot(aes(gamma)) +
  geom_density(fill="blue",alpha=.5,color="blue") 
},
filename = "../figs/gamma_dist.png",
       width = 8, height = 6, device = "png", dpi=600)
```

filter documents with low posterior
```{r}
btw.reg <- btw %>%
  filter(gamma >= 0.5) %>%
  #filter(!site=="stern.de") %>%
  mutate(topic = factor(topic),
         site = factor(site),
         month = factor(month),
         fb_shares_log = log1p(fb_shares)) -> btw.reg
```

Proportion of zeros
```{r}
nrow(btw.reg[which(btw.reg$fb_shares==0),])/nrow(btw.reg)
```

Distribution of Facebook shares
```{r}
btw.reg$share.cut <- cut(btw.reg$fb_shares, 
                         breaks = c(0,1,100,500,1000,5000,max(btw.reg$fb_shares)),
                         right = T, include.lowest = TRUE)

ggsave(plot={
  ggplot(btw.reg, aes(x=share.cut)) + 
  geom_bar(aes(fill=site), alpha=.7) + 
  labs(title = "") +
  labs(x="Facebook Shares", y="Count", fill="") +
  scale_fill_manual(values = col) +
  scale_x_discrete(labels = c("0","1-100","101-500","501-1000","1001-5000",">5000"))
},
filename = "../figs/facebook_shares.png", width = 8, height = 6, 
       device = "png", dpi = 600
)
```

## Topic Timeline
```{r Data for topic trends chart, message=FALSE, warning=FALSE}
set.seed(7292)
btw.reg %>% 
  mutate(date = as.POSIXct(date),
         allocation = 1) %>%
  mutate(post_period = 
           date %>% with_tz("Europe/Paris") %>% 
           floor_date("week")) %>%
  # Summarise into tidy dataframe
  group_by(post_period, topic, topic_name) %>% dplyr::summarise(articles = sum(allocation)) %>%
  # Mark peaks 
  group_by(topic_name) %>% dplyr::mutate(peak = ifelse(articles==max(articles),1,0)) -> chart_topic_trends.dt
```

```{r message=FALSE, warning=FALSE}
ggsave(
  plot = {
    ggplot(chart_topic_trends.dt, 
       aes(post_period, articles, color=topic_name)) +
  geom_line() +
  scale_color_manual(
            values = rainbow(28) %>% 
              adjustcolor(red.f = 0.6, green.f = 0.6, blue.f = 0.6)
          ) +
  geom_text_repel(data=subset(chart_topic_trends.dt,peak==1),
            aes(post_period,articles,label=paste(topic,topic_name,sep=":")),
            size=2.8) +
      geom_vline(aes(xintercept=as.POSIXct("2017-09-23")),
                 linetype = 2, color="grey20") +
      geom_vline(aes(xintercept=as.POSIXct("2017-11-19")),
                 linetype = 2, color="grey20") +
  ylim(c(0,200)) +
    theme(
      legend.position   = "none",
      #plot.background   = element_rect("#fafafa", "#fafafa"),
      plot.title        = element_text(size = 18),
      plot.subtitle     = element_text(size = 8),
      plot.caption      = element_text(size = 7),
      panel.grid        = element_blank(),
      #panel.background  = element_rect("#fafafa"),
      axis.title.x      = element_blank(),
      axis.text.y       = element_text(size = 8)
    )
  },
  device = "png",
  filename = "../figs/topic-timeline.png",
  dpi = 600, height = 5, width = 8
)
```

## Analysis
```{r}
summary(btw.reg[,c("site","text_length","fb_shares","gamma")])
```

### Simple linear regression
```{r}
source("func/functions.R")

library(stargazer)
lm <- lm(fb_shares_log ~ topic + site + text_length, data=btw.reg)

lm.t <- stargazer(lm)

lm.t <- AddResizeBox(lm.t)

writeLines(lm.t, con = "../doc/lm1.tex")
```

### Zero-inflated Poisson regression
The distribution of the outcome can then be modeled in terms of two parameters, $\pi$ the probability of 'always zero', and $\mu$, the mean number of publications for those not in the 'always zero' group. A natural way to introduce covariates is to model the logit of the probability $\pi$ of always zero and the log of the mean $\mu$ for those not in the always zero class.

```{r Zero-inflated Poisson}
library(pscl)
m1 <- zeroinfl(fb_shares ~ topic + site | topic + site + text_length, data = btw.reg)
summary(m1)
```
Looking at the inflate equation we see that some topics are significant predictors of being in the 'always zero' class.

@@

```{r}
p1 <- glm(fb_shares ~ topic, family = poisson, data = btw.reg)
summary(p1)
```

@@

The predicted values are plotted in black. They are connected to the actual data points through a line and high residuals (in abolsute terms) are made more red on actual values. This helps to identify non-linearity in the data. We can see that there is more red for extreme values for spiegel.de as well as for the topics 12, 21 and 9. 
```{r message=FALSE, warning=FALSE}
btw.reg %>%
  select(site, fb_shares, topic, residuals, predicted) %>%
  gather(key="iv", value="x", -fb_shares, -predicted, -residuals) %>%
  ggplot(aes(x=x, y=fb_shares)) +
  geom_segment(aes(xend=x, yend=predicted), alpha=.2) +
  geom_point(aes(color=residuals)) +
  scale_color_gradient2(low="black", mid="blue", high="red") +
  guides(color=FALSE) +
  geom_point(aes(y=predicted), shape=1) +
  facet_grid(~iv, scales="free") +
  theme_bw()
```
